---
title: "fitbir"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(rlang)
library(ggplot2)
library(dplyr)
library(psych)

all_fitbir <- read_excel("/Users/emmatinney/Library/Mobile Documents/com~apple~CloudDocs/Documents/Northeastern/CCBH_Lab/Projects/fitbir/track_tbi.xlsx")
dti_fitbir <- read_excel("/Users/emmatinney/Library/Mobile Documents/com~apple~CloudDocs/Documents/Northeastern/CCBH_Lab/Projects/fitbir/fitbir_dti.xlsx")
df_fitbir <- left_join(dti_fitbir, all_fitbir, by = "id")
df_fitbir$gose_baseline <- as.numeric(df_fitbir$gose_baseline)
df_fitbir$gose_three <- as.numeric(df_fitbir$gose_three)
df_fitbir$gose_six <- as.numeric(df_fitbir$gose_six)
df_fitbir$SES <- as.numeric(df_fitbir$SES)
df_fitbir$gcs_severity <- as.numeric(df_fitbir$gcs_severity)

df_fitbir$ISI_total_score_base <- as.numeric(df_fitbir$ISI_total_score_base)
df_fitbir$ISI_total_score_three <- as.numeric(df_fitbir$ISI_total_score_three)
df_fitbir$ISI_total_score_six <- as.numeric(df_fitbir$ISI_total_score_six)
df_fitbir$ISI_total_score_twelve <- as.numeric(df_fitbir$ISI_total_score_twelve)
df_fitbir$gose_twelve_dicho <- as.numeric(df_fitbir$gose_twelve_dicho)
df_fitbir$TMTPartATime_two <- as.numeric(df_fitbir$TMTPartATime_two)
df_fitbir$TMTPartATime_six <- as.numeric(df_fitbir$TMTPartATime_six)
df_fitbir$TMTPartATime_twelve <- as.numeric(df_fitbir$TMTPartATime_twelve)
df_fitbir$TMTPartBTime_two <- as.numeric(df_fitbir$TMTPartBTime_two)
df_fitbir$TMTPartBTime_six <- as.numeric(df_fitbir$TMTPartBTime_six)
df_fitbir$TMTPartBTime_twelve <- as.numeric(df_fitbir$TMTPartBTime_twelve)
df_fitbir$RAVLTsum_two <- as.numeric(df_fitbir$RAVLTsum_two)
df_fitbir$RAVLTlearnt2 <- as.numeric(df_fitbir$RAVLTlearnt2)    
df_fitbir$RAVLTsum_6m <- as.numeric(df_fitbir$RAVLTsum_6m) 
df_fitbir$RAVLTlearnt_6m <- as.numeric(df_fitbir$RAVLTlearnt_6m) 
df_fitbir$RAVLTsum_12m <- as.numeric(df_fitbir$RAVLTsum_12m) 
df_fitbir$RAVLTlearnt_12m <- as.numeric(df_fitbir$RAVLTlearnt_12m) 
df_fitbir$educ <- as.numeric(df_fitbir$educ) 
df_fitbir$SES <- as.numeric(df_fitbir$SES) 
df_fitbir$age <- as.numeric(df_fitbir$age) 
df_fitbir$WAISscore_two <- as.numeric(df_fitbir$WAISscore_two) 
df_fitbir$WAISscore_six <- as.numeric(df_fitbir$WAISscore_six) 
df_fitbir$WAISscore_twelve <- as.numeric(df_fitbir$WAISscore_twelve) 
df_fitbir$PCS12_base <- as.numeric(df_fitbir$PCS12_base) 
df_fitbir$PCS12_three <- as.numeric(df_fitbir$PCS12_three) 
df_fitbir$PCS12_six <- as.numeric(df_fitbir$PCS12_six) 
df_fitbir$PCS12_twelve <- as.numeric(df_fitbir$PCS12_twelve) 

df_fitbir$MCS12_base <- as.numeric(df_fitbir$MCS12_base) 
df_fitbir$MCS12_three <- as.numeric(df_fitbir$MCS12_three) 
df_fitbir$MCS12_six <- as.numeric(df_fitbir$MCS12_six) 
df_fitbir$MCS12_twelve <- as.numeric(df_fitbir$MCS12_twelve) 

x <- as.factor(df_fitbir$site_name)
levels(x)
df_fitbir$site_numeric <- df_fitbir$site_name
df_fitbir$site_numeric[x == "BCM" | x == "Baylor College of Medicine"] <- 1
df_fitbir$site_numeric[x == "VCU"] <- 11
df_fitbir$site_numeric[x == "The Univ. of Texas at Austin IRC"] <- 8
df_fitbir$site_numeric[x == "UCSF QB3"] <- 3
df_fitbir$site_numeric[x == "MRRC"] <- 7
df_fitbir$site_numeric[x == "MGH"] <- 2
df_fitbir$site_numeric[x == "UC Radiology 3T"] <- 4
df_fitbir$site_numeric[x == "University of Maryland Medical System"] <- 5
df_fitbir$site_numeric[x == "Joseph Applebaum Center"] <- 6
df_fitbir$site_numeric[x == "SIEMENS 3"] <- 5
df_fitbir$site_numeric[x == "UPMC HEALTH SYS PRESBY" | x == "UPMC Presbyterian University Hospital,"] <- 7
df_fitbir$site_numeric[x == "MEADOWS MRI"] <- 9
df_fitbir$site_numeric[x == "University of Washington Medical Center" | x == "University of Washington"] <- 10
df_fitbir$site_numeric[x == "CIND"] <- 3
df_fitbir$site_numeric[x == "UT SOUTHWESTERN MEDICAL CENTER"] <- 9
df_fitbir$site_numeric[x == "Univ of Texas at Austin"] <- 8

df_fitbir$site_numeric <- as.numeric(df_fitbir$site_numeric)

x <- as.factor(df_fitbir$case_control)
levels(x)
df_fitbir$tbi <- df_fitbir$case_control
df_fitbir$tbi[x == "Case"] <- 1
df_fitbir$tbi[x == "Control"] <- 0
df_fitbir$tbi <- as.numeric(df_fitbir$tbi)

case <- subset(df_fitbir, tbi=="1")
control <- subset(df_fitbir, tbi=="0")
#remove the subjects without any label
df_filtered <- df_fitbir[complete.cases(df_fitbir$tbi), ]
df_fitbir$anxiety[is.na(df_fitbir$anxiety)] = 0
df_fitbir$depression[is.na(df_fitbir$depression)] = 0
df_fitbir$diabetes_typeI[is.na(df_fitbir$diabetes_typeI)] = 0
df_fitbir$diabetes_typeII[is.na(df_fitbir$diabetes_typeII)] = 0
df_fitbir$develop_dis[is.na(df_fitbir$develop_dis)] = 0
df_fitbir$hypertension[is.na(df_fitbir$hypertension)] = 0
```


```{r}
pairs.panels(df_fitbir[,c('age', 'sex', 'educ', 'SES')])
```

#19=left anterior internal capsule
```{r}
#difference between case vs control
summary(lm(df_fitbir$AD_19 ~ df_fitbir$tbi + df_fitbir$SES + df_fitbir$educ + df_fitbir$age  + df_fitbir$sex +  df_fitbir$site_numeric))

#interaction between case vs control and ISI
summary(lm(df_fitbir$AD_19 ~ df_fitbir$ISI_total_score_base + df_fitbir$tbi:df_fitbir$ISI_total_score_base + df_fitbir$tbi + df_fitbir$SES + df_fitbir$educ + df_fitbir$age + df_fitbir$sex +  df_fitbir$site_numeric))

summary(lm(df_fitbir$AD_19 ~ df_fitbir$ISI_total_score_three + df_fitbir$tbi:df_fitbir$ISI_total_score_three + df_fitbir$tbi + df_fitbir$SES + df_fitbir$educ + df_fitbir$age + df_fitbir$sex +  df_fitbir$site_numeric))

summary(lm(df_fitbir$AD_19 ~ df_fitbir$ISI_total_score_six + df_fitbir$tbi:df_fitbir$ISI_total_score_six + df_fitbir$tbi + df_fitbir$SES + df_fitbir$educ + df_fitbir$age + df_fitbir$sex +  df_fitbir$site_numeric))

summary(lm(df_fitbir$AD_19 ~ df_fitbir$ISI_total_score_twelve + df_fitbir$tbi:df_fitbir$ISI_total_score_twelve + df_fitbir$tbi + df_fitbir$SES + df_fitbir$educ + df_fitbir$age + df_fitbir$sex +  df_fitbir$site_numeric))

#correlation between CASES AD AIC_L and ISI
summary(lm(case$AD_19 ~ case$ISI_total_score_base + case$SES + case$educ + case$age + case$sex  +  case$site_numeric))
summary(lm(case$AD_19 ~ case$ISI_total_score_three + case$SES + case$educ + case$age + case$sex  +  case$site_numeric))
summary(lm(case$AD_19 ~ case$ISI_total_score_six + case$SES + case$educ + case$age + case$sex  +  case$site_numeric))
summary(lm(case$AD_19 ~ case$ISI_total_score_twelve + case$SES + case$educ + case$age + case$sex  +  case$site_numeric))

#correlation between CONTROLS AD AIC_L and ISI
summary(lm(control$AD_19 ~ control$ISI_total_score_base + control$SES + control$educ + control$age + control$sex  +  control$site_numeric))
summary(lm(control$AD_19 ~ control$ISI_total_score_three + control$SES + control$educ + control$age + control$sex +  control$site_numeric))
summary(lm(control$AD_19 ~ control$ISI_total_score_six + control$SES + control$educ + control$age + control$sex +  control$site_numeric))
summary(lm(control$AD_19 ~ control$ISI_total_score_twelve + control$SES + control$educ + control$age + control$sex +  control$site_numeric))


#correlation between cases AD AIC_L and ravlt
summary(lm(case$AD_19 ~ case$RAVLTlearnt2 + case$SES + case$educ + case$age + case$sex +  case$site_numeric))
summary(lm(case$AD_19 ~ case$RAVLTlearnt_6m + case$SES + case$educ + case$age + case$sex +  case$site_numeric))
summary(lm(case$AD_19 ~ case$RAVLTlearnt_12m + case$SES + case$educ + case$age + case$sex +  case$site_numeric))

```
```
```{r}
mean(case$AD_19) 
mean(control$AD_19)
sd(case$AD_19) 
sd(control$AD_19)
mean(case$ISI_total_score_base,  na.rm = TRUE)
mean(case$ISI_total_score_three,  na.rm = TRUE) 
mean(case$ISI_total_score_six,  na.rm = TRUE) 
mean(case$ISI_total_score_twelve,  na.rm = TRUE) 
mean(control$ISI_total_score_twelve,  na.rm = TRUE) 
mean(control$ISI_total_score_six,  na.rm = TRUE) 
mean(control$ISI_total_score_three,  na.rm = TRUE)
mean(control$ISI_total_score_base,  na.rm = TRUE) 

sd(case$ISI_total_score_base,  na.rm = TRUE)
sd(case$ISI_total_score_three,  na.rm = TRUE) 
sd(case$ISI_total_score_six,  na.rm = TRUE) 
sd(case$ISI_total_score_twelve,  na.rm = TRUE) 
sd(control$ISI_total_score_twelve,  na.rm = TRUE) 
sd(control$ISI_total_score_six,  na.rm = TRUE) 
sd(control$ISI_total_score_three,  na.rm = TRUE)
sd(control$ISI_total_score_base,  na.rm = TRUE) 

mean(case$RAVLTlearnt2,  na.rm = TRUE)
mean(case$RAVLTlearnt_6m,  na.rm = TRUE) 
mean(case$RAVLTlearnt_12m,  na.rm = TRUE) 
mean(control$RAVLTlearnt2,  na.rm = TRUE) 
mean(control$RAVLTlearnt_6m,  na.rm = TRUE) 
mean(control$RAVLTlearnt_12m,  na.rm = TRUE)
mean(case$gcs_severity,  na.rm = TRUE)

sd(case$RAVLTlearnt2,  na.rm = TRUE)
sd(case$RAVLTlearnt_6m,  na.rm = TRUE) 
sd(case$RAVLTlearnt_12m,  na.rm = TRUE) 

t.test(control$ISI_total_score_base, case$ISI_total_score_base)
t.test(control$ISI_total_score_three, case$ISI_total_score_three)
t.test(control$ISI_total_score_six, case$ISI_total_score_six)
t.test(control$ISI_total_score_twelve, case$ISI_total_score_twelve)
t.test(control$RAVLTlearnt2, case$RAVLTlearnt2)
t.test(control$RAVLTlearnt6, case$RAVLTlearnt6)
t.test(control$RAVLTlearnt12, case$RAVLTlearnt12)

mean(case$SES,  na.rm = TRUE)
mean(control$SES,  na.rm = TRUE)
mean(case$age,  na.rm = TRUE) 
mean(control$age,  na.rm = TRUE)
mean(case$educ,  na.rm = TRUE)
mean(control$educ,  na.rm = TRUE)

sd(case$SES,  na.rm = TRUE)
sd(control$SES,  na.rm = TRUE)
sd(case$age,  na.rm = TRUE) 
sd(control$age,  na.rm = TRUE)
sd(case$educ,  na.rm = TRUE)
sd(control$educ,  na.rm = TRUE)

t.test(control$age, case$age)
t.test(control$educ, case$educ)
t.test(control$SES, case$SES)
#test of proportions
count(control, sex)
count(case, sex)
prop.test(x = c(28, 133), n = c(85, 374))

```




```{r, echo=FALSE}
library(rlang)
library(ggplot2)
scatterplot <- function(dataframe, x, y, ylab, xlab) {
  ggplot(data = dataframe, aes(x={{x}}, y = {{y}})) + 
    geom_point(color =  "#663399") + 
    theme(
      # Hide panel borders and remove grid lines
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      # Change axis line
      axis.line = element_line(colour = "black")
    ) +
    labs(y=ylab, x = xlab) +
    
    stat_smooth(method = "lm", col = "#cc8400")
}

library(ggplot2)
# a basic plot to plot linear regression results, colour coding the dots and lines
ggplot(fitcog_55, aes(x = fitcog_55$crf_log, y = fitcog_55$reasoning_VS_PCA)) + 
  geom_point(color =  "#663399") + 
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black")
  ) +
  labs(y="reasoning", x = "V02peak_log") +
  stat_smooth(method = "lm", col = "#cc8400")

library(ggplot2)
TBI$AVPOS_change <- as.numeric(TBI$AVPOS_change)
# a basic plot to plot linear regression results, colour coding the dots and lines
ggplot(case, aes(x = AD_19, y = RAVLTlearnt_12m)) + 
  geom_point(color = "#d95f0e") + 
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black")
  ) +
  labs(y="RAVLT Delayed Recall at 12 months post injury", x = "Axial Diffusivity") +
  stat_smooth(method = "lm", col = "#d95f0e")

#Same as plot above but for interaciton effects (colour codes dots per a categorical varibale and plots separate lines)
library(ggplot2)
ggplot(df_filtered, aes(x = AD_19, y = ISI_total_score_twelve)) + 
  geom_point(aes(color = factor(case_control))) + 
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black")
  ) + theme(text = element_text(size = 18), plot.title = element_text(size = 20)) +
  labs(y="ISI score at 12 months", x = "Axial Diffusivity") +
  stat_smooth(method = "lm", aes(color = factor(case_control))) +
  scale_color_manual(values = c("#d95f0e", "#2c7fb8"), 
                     labels = c("TBI", "No TBI"))

```


```{r, echo=FALSE}
ggplot(data_adni, aes(x = PSQI, y = AD_anterior_internal_cap_l)) +
  geom_point(aes(color = factor(TBI))) +
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black")
  ) +   theme(text = element_text(size = 20), plot.title = element_text(size = 20)) +
  labs(y = "Axial Diffusivity", x = "Sleep Disturbances") +
  stat_smooth(method = "lm", aes(color = factor(TBI))) +
  scale_color_manual(values = c("#d95f0e", "#2c7fb8"), 
                     labels = c("No TBI", "TBI"))

# A box plot for two groups based  
library(ggplot2)
ggplot(df_filtered, aes(group=case_control, x=case_control, y=AD_19, case_control)) +
  geom_point(aes(color = factor(case_control))) + 
  geom_boxplot(aes(color = factor(case_control))) + 
  theme_classic() + 
  theme(text = element_text(size = 18), plot.title = element_text(size = 20)) +
  labs(y="Axial Diffusivity in AIC (L)", x = "Group") +
  scale_color_manual(values = c("#d95f0e", "#2c7fb8"), 
                     labels = c("TBI", "No TBI"))

ggplot(mtcars, aes(x = mpg, y = drat)) +
  geom_point(color = "56B4E9")

geom_point(aes(color = factor(V02max$Sexo))) #for cat split 

# boxx plots for means of n=x groups 
library("ggpubr")
ggboxplot(IPL_long_mid, x = "scout", y = "current", 
          color = "scout", palette = c("#81d8d0", "#1e90ff", "#ff8c00", "#800080", "#ffbf00", "#808080"),
          ylab = "current", xlab = "scout")

#order = c("ctrl", "trt1", "trt2"),
#c("#00AFBB", "#E7B800", "#FC4E07", ),

#cMf  = #81d8d0
#mPFC = #1e90ff
#sFr = #ff8c00
#m1 = #808080
#IPL = #ffbf00
#sPa= #800080

# colour palletes and color codes 

library(wesanderson)
pal <- wes_palette("Darjeeling1")
  
#skyish blue "#56B4E9"
#Orange "#E69F00"
#reddish #d21e2d
#teal "#1ed2c3"
#pink #d21e87
#green #1ed269
#rebecca purple #663399
#dirty orange #cc8400
# red #d21e2d
#BR green #007542
# burgendy #7500
# green/blue #81d8d0


```

